OneD: increasing reproducibility of Hi-C Samples with abnormal karyotypes

### 2017-06-07 main.tex
Verb missing in line 98
Most changes
Comments

- lines 104 and 107: what do you mean by explicitily and implicitily, respectively?
- line 192: what do you mean by "the content"?
- lines 358-360: there seems to be mixed text pieces, and it converts wrong to the PDF
- F. Le Dily told me that that the ERC/Synergy has to be acknowledged as follows: "We received funding from the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013)/ERC Synergy grant agreement 609989 (4DGenome). The content of this manuscript reflects only the author’s views and the Union i

Contrary to the precursor technologies 3C, 4C and 5C \citep{de2012decade},
Hi-C interrogates all possible pairwise interactions between restriction
fragments. However, this does not guarantee that the method has no bias.
On the contrary, local genome features such as the G+C content, the
availability of restriction enzyme sites and the mappability of the
sequencing reads have been shown to impact the results
\citep{yaffe2011probabilistic}. In addition, general experimental biases
such as batch effects. It is thus important to normalize Hi-C data in
order to remove biases and artifacts, so that they are not confused with
biological signal.

@@

Neither of these strategies were designed with regard for cell types with
karyotypic aberrations, most common in cancer cells. Hi-C is very
sensitive to aneuploidy, copy number variations and translocations.
Actually, these aberrations have so much influence on the outcome that the
artifacts can be used to re-assemble the target genome
\citep{korbel2013genome}. So far the only attempt to address the issue
was the chromosome-adjusted Iterative Correction Bias method
(\textit{caICB}) of \cite{wu2016computational}.  However, caICB applies a
uniform chromosome-wide correction, effectively excluding the numerous
cases of partial aneuploidy and regional copy number variations.

@@

The most common representation of Hi-C data is a contact matrix, obtained
by slicing the genome in $n$ consecutive bins of fixed size (the
resolution) and computing the number of contacts between each pair of
bins. The values are stored in the cells of the contact matrix ($x_{ij}$),
quantifying of the interaction between the tow loci at positions $i$ and
$j$.

@@

All data were processed though a pipeline based on TADBit
\citep{serra2016structural}. Briefly, after controlling the quality of
fastq files, paired-end reads were mapped in the corresponding reference
genome (hg38 and mm10) taking into account the restriction enzyme site.
Non-informative contacts were removed applying the following TADBit
filters: self-circle, dangling-ends, error, extra dangling-end, duplicated
and random-breaks. For more details, see the methods section of
\cite{stadhouders2017transcription}. In addition, the pipeline is
available from the supplementary material published by
\cite{quilez2017managing}.

@@

\begin{figure}[b]
	\centerline{\includegraphics[width=.45\textwidth]
{img/copy_number_figure2.pdf}}
\caption{Pearson correlation between the total number of contacts per bin
and an independent estimation of the copy number (COSMIC). Left panel
T47D breast cancer cell line, right panel K562 leukemia cell line.}
\label{fig:copy_number}
\end{figure}

@@

Using the other metrics described in \ref{sec:comp} yielded similar
results (Supplementary Figures 5, 6 and 7). Note that Spearman correlation
of contacts presented the worst performance for all scenarios, and it thus
seems to be a poor choice as a metric to compare HiC matrices.

@@

\textit{OneD} is the second fastest method and it always ran in less than
1 min in the conditions of the benchmark. We occasionally encountered
problems where \textit{ICE} ran faster than \textit{OneD} so it is fair to
say that none of them has a definite speed advantage. Throughout this
benchmark, the memory footprint of \textit{OneD} was less than 300 MB.
Interestingly, the running time of \textit{OneD} is much more homogeneous
than that of the other methods. The reason is that the size of the
regression problem to be solved by the \texttt{mgcv} package is always the
same at fixed resolution.

@@

This raises the question whether variations of the copy number constitute
a biological signal or an artifact. If the biological sample contains
karytoypic aberrations, then its genome is grossly different the reference
so there can be no correct way to normalize the Hi-C signal. The proper
approach would be to use the actual genome of the biological sample as a
reference to construct the contact map. However, this strategy is
presently unfeasible because assembling mammalian genomes is still a hard
problem.

>>>>

Contrary to the precursor technologies 3C, 4C and 5C \citep{de2012decade},
Hi-C interrogates all possible pairwise interactions between restriction
fragments. However, this does not guarantee that the method has no bias.
On the contrary, local genome features such as the G+C content, the
availability of restriction enzyme sites and the mappability of the
sequencing reads have been shown to impact the results
\citep{yaffe2011probabilistic}. In addition, general experimental biases
such as batch effects exist. It is thus important to normalize Hi-C data in
order to remove biases and artifacts, so that they are not confused with
biological signal.

@@

Neither of these strategies were designed for cell types with
karyotypic aberrations, most common in cancer cells. Hi-C is very
sensitive to aneuploidy, copy number variations and translocations.
Actually, these aberrations have so much influence on the outcome that the
artifacts can be used to re-assemble the target genome
\citep{korbel2013genome}. So far the only attempt to address the issue
was the chromosome-adjusted Iterative Correction Bias method
(\textit{caICB}) of \cite{wu2016computational}.  However, caICB applies a
uniform chromosome-wide correction, effectively excluding the numerous
cases of partial aneuploidy and regional copy number variations.

@@

The most common representation of Hi-C data is a contact matrix, obtained
by slicing the genome in $n$ consecutive bins of fixed size (the
resolution) and computing the number of contacts between each pair of
bins. The values are stored in the cells of the contact matrix ($x_{ij}$),
quantifying of the interaction between the two loci at positions $i$ and
$j$.

@@

All data were processed through a pipeline based on TADbit
\citep{serra2016structural}. Briefly, after controlling the quality of
FASTQ files, paired-end reads were mapped to the corresponding reference
genome (hg38 and mm10) taking into account the restriction enzyme site.
Non-informative contacts were removed applying the following TADbit
filters: self-circle, dangling-ends, error, extra dangling-end, duplicated
and random-breaks. For more details, see the methods section of
\cite{stadhouders2017transcription}. In addition, the pipeline is
available from the supplementary material published by
\cite{quilez2017managing}.

@@

\begin{figure}[b]
	\centerline{\includegraphics[width=.45\textwidth]
{img/copy_number_figure2.pdf}}
\caption{Pearson correlation between the corrected number of contacts per bin
and an independent estimation of the copy number (COSMIC). Left panel
T47D breast cancer cell line, right panel K562 leukemia cell line.}
\label{fig:copy_number}
\end{figure}

@@

Using the other metrics described in \ref{sec:comp} yielded similar
results (Supplementary Figures 5, 6 and 7). Note that Spearman correlation
of contacts presented the worst performance for all scenarios, and it thus
seems to be a poor choice as a metric to compare Hi-C matrices.

@@

\textit{OneD} is the second fastest method and it always ran in less than
1 min in the conditions of the benchmark. We occasionally encountered
samples where \textit{ICE} ran faster than \textit{OneD} so it is fair to
say that none of them has a definite speed advantage. Throughout this
benchmark, the memory footprint of \textit{OneD} was less than 300 MB.
Interestingly, the running time of \textit{OneD} is much more homogeneous
than that of the other methods. The reason is that the size of the
regression problem to be solved by the \texttt{mgcv} package is always the
same at fixed resolution.

@@

This raises the question whether variations of the copy number constitute
a biological signal or an artifact. If the biological sample contains
karytoypic aberrations, then its genome is grossly different from the reference
so there can be no correct way to normalize the Hi-C signal. The proper
approach would be to use the actual genome of the biological sample as a
reference to construct the contact map. However, this strategy is
presently unfeasible because assembling mammalian genomes is still a hard
problem.
